\Large Exponential Random Graph Models with Big Networks: Maximum Pseudolikelihood Estimation and the\\Parametric Bootstrap

### 2017-06-27 TeX/paper.tex
Add files via upload

Alternating k-star                              & 0.125                            & 0.064                             &     0.108                             & 0.006                                   & -0.011 & 0.2379  \\ \hline

@@

One of the major disadvantages of MPLE over MCMLE is that it cannot assess {\it degeneracy}. When fitting an ERGM one usually has to deal with the problem of unreliable approximatively likelihood estimates for the model's parameters. The reason why degeneracy occurs is that the stochastic process generated by the MCMC-algorithm does not necessarily hold through the model's defined distribution of the random variable $Y$ as stationary distribution (see Handcock \cite{Handcock.2003} and Rinaldo et al. \cite{Rinaldo2009}). The bootstrapped MPLE, however, allows assessing degenerate models as well. In order to verify whether a model is degenerate or not, one can take a look at density and trace plots as visualized in figure ??. The trace plots on the left side depict the the attained values via MCMC simulated networks for every single statistic included into the model, centered on the statistic values of the observed network. The plots on the right side visualize the empirical density function of the respective statistic, based on the simulated networks (Hunter and Handcock \cite{Hunter.2006}). For a non-degenerated model the empirical density function should be symmetrical around zero for every included centered statistic, since this corresponds with the expected value of a centered statistic (compare equation (\ref{assumption})). 

@@

$$\text{insert density plots and trace plots}$$

\begin{itemize}

\item describe advantage over MPLE (better coverage) and MCMLE (faster)
\item can easily be parallelized/ using multiple cores -> increases speed
\item this method is also able to detect degeneracy
\item include table with boot MPLE results and MCMLE results
\end{itemize}

>>>>

Alternating k-star                              & 0.124                            & 0.064                             &     0.108                             & 0.006                                   & -0.011 & 0.2379  \\ \hline

@@

One of the major disadvantages of MPLE over MCMLE is that it cannot assess {\it degeneracy}. When fitting an ERGM one usually has to deal with the problem of unreliable approximatively likelihood estimates for the model's parameters. The reason why degeneracy occurs is that the stochastic process generated by the MCMC-algorithm does not necessarily hold through the model's defined distribution of the random variable $Y$ as stationary distribution (see Handcock \cite{Handcock.2003} and Rinaldo et al. \cite{Rinaldo2009}). The bootstrapped MPLE, however, allows assessing degenerate models as well. In order to verify whether a model is degenerate or not, one can take a look at density and trace plots as visualized in figure \ref{diagnostics}. The trace plots on the left side depict the the attained values via MCMC simulated networks for every single statistic included into the model, centered on the statistic values of the observed network. The plots on the right side visualize the empirical density function of the respective statistic, based on the simulated networks (Hunter and Handcock \cite{Hunter.2006}). For a non-degenerated model the empirical density function should be symmetrical around zero for every included centered statistic, since this corresponds with the expected value of a centered statistic (compare equation (\ref{assumption})). 

@@

\includegraphics[scale=.6]{MPLE_histogram_traceplot_cosponsorship}
\caption{Network statistics of the 500 boootstrap samples for the cosponsorship network. The thick line in both, the traceplots and the histograms, represents the network statistics of the observed network. }
\label{diagnostics}
